---
title: "Installation"
format: rst
date: 2023-04-12
number-sections: true
---

## Installing Giotto Suite
Giotto Suite is installed via R **but there are required Python modules** which **must** be installed in order for Giotto Suite to function properly.
Below are the instructions for both the installation of the Giotto Suite package and required Python environment.

### Requirements
* R (>= 3.5.1)
* Python (>= 3.0)
* `Windows, MacOS or Linux specific installation tools for `Posit <https://support.posit.co/hc/en-us/articles/200486498-Package-Development-Prerequisites>`_ (previously RStudio)

## Installation

### Install Giotto Suite
```{r eval=FALSE, message=FALSE}
# Necessary for installation from R
if(!"devtools" %in% installed.packages()) {
  install.packages("devtools")
}

devtools::install_github("drieslab/Giotto@suite")

```

### Install the Giotto Python Environment
```{r eval=FALSE}
library(Giotto)
installGiottoEnvironment()
```

## Try Giotto in the cloud
You can also run analyses in Giotto on Terra.bio. Take a look on the `Terra tutorial <https://giottosuite.readthedocs.io/en/latest/subsections/trygiotto/terra.html#terra/>`_.


Encountering errors? Checkout out the `Troubleshooting page <../errorsfaqsandtips.html>`__ for help.
Alternatively, post to an issue to the Giotto GitHub page `here <https://github.com/drieslab/Giotto>`__.
Please include the version numbers of R, Giotto, and the OS in use at the time of the issue.

```{r eval=TRUE, include=FALSE}
library(Giotto)
```


## Session Info
```{r}
sessionInfo()
```
