---
title: "Saving Plots"
format: rst
date: 2023-04-14
number-sections: true
---

## Explanation
This tutorial walks through saving spatial data in plots.
*Please see the [Configuration](./configuration_04122023.html) and [Giotto Object](./getting_started_gobject.html) vignettes before walking through this tutorial.*

R/Posit and Giotto provide different ways to save spatial data. 
Here, a **giottoObject** will be created without using **giottoInstructions** so that the save parameters for plotting functions within Giotto as well as the default saving methods built into R/Posit may be emphasized here. 
Note that for plotting functions, all parameters available to the *save_param* argument may be found by running **showSaveParameters()**.

## Start Giotto
```{r eval=TRUE, message=FALSE}
# Ensure Giotto Suite is installed
if(!"Giotto" %in% installed.packages()) {
  devtools::install_github("drieslab/Giotto@suite")
}
library(Giotto)

# Ensure Giotto Data is installed
if(!"GiottoData" %in% installed.packages()) {
  devtools::install_github("drieslab/GiottoData")
}
library(GiottoData)

# Ensure the Python environment for Giotto has been installed
genv_exists = checkGiottoEnvironment()
if(!genv_exists){
  # The following command need only be run once to install the Giotto environment
  installGiottoEnvironment()
}
```

## Create a Giotto object
Since the focus of this vignette is saving methods, the **giottoObject** will not be created with **giottoInstructions**. See [Giotto Object](./getting_started_gobject.html) for further intuition on working with a **giottoObject** that has been provided instructions. 

```{r eval=TRUE}
data_directory = paste0(getwd(),'/')

# Download dataset 
getSpatialDataset(dataset = 'osmfish_SS_cortex', directory = data_directory, method = 'wget')

# Specify path to files
osm_exprs = paste0(data_directory, "osmFISH_prep_expression.txt")
osm_locs = paste0(data_directory, "osmFISH_prep_cell_coordinates.txt")
meta_path = paste0(data_directory, "osmFISH_prep_cell_metadata.txt")

## CREATE GIOTTO OBJECT with expression data and location data
my_gobject <- createGiottoObject(expression = osm_exprs,
                                 spatial_locs = osm_locs)

metadata = data.table::fread(file = meta_path)
my_gobject = addCellMetadata(my_gobject, new_metadata = metadata,
                             by_column = T, column_cell_ID = 'CellID')
```

## Standard R Save Methods

Note that by default, plotting functions will return a plot object that may be saved or further manipulated.
Manually save plot as a PDF in the current working directory:

```{r eval=TRUE}
save_path = paste0(getwd(),'/first_plot.pdf')
library(ggplot2)
# This function serves only to ensure the following lines run consecutively.
save_pdf_plot <- function(){ 
  pdf(file = save_path, width = 7, height = 7)
  pl = spatPlot(my_gobject)
  dev.off()
}

save_pdf_plot()

### Plot clusters, edit plot object, then save using the ggplot add-on, cowplot:
mypl = spatPlot(gobject = my_gobject, 
                cell_color = 'ClusterName')

# Add a black background
mypl = mypl + theme(panel.background = element_rect(fill ='black'),
                    panel.grid = element_blank())

# Add a legend
mypl = mypl + guides(fill = guide_legend(override.aes = list(size=5)))

# Save in the current working directory
cowplot::save_plot(plot = mypl,
                   filename = 'clusters_black.png', 
                   path = getwd(),
                   device = png(),
                   dpi = 300, 
                   base_height = 10, 
                   base_width = 10)
```

## Save Plot Directly to the Default Folder
The default save folder is the current working directory. This will be the case if instructions are not provided, or if a *save_dir* is not specified within **giottoInstructions**. See the [**createGiottoInstructions** documentation](http://giottosuite.com/reference/createGiottoInstructions.html) and [Giotto Object](./getting_started_gobject.html) for default arguments and more details.
```{r eval=TRUE}
spatPlot(my_gobject, 
         cell_color = 'ClusterName', 
         save_plot = TRUE)
```


##  Save plot Directly, but Overwrite Default Save Parameters

In this example, assume it is desired that the plot is: 
- Shown in the console
- Not returned as an object from the plotting function call 
- Saved in a subdirectory of the current working directory as a .png file with a dpi of 200, height of 9 inches, and width of 9 inches.
- Saved with the file name "my_name"

See [Giotto Object](./getting_started_gobject.html) for more details.  
Run the command **showSaveParameters()** to see all available parameters.   

```{r eval=TRUE}
# Specify new subdirectory name
results_directory =  'my_subfolder/'

# Plot clusters, create, and save to a new subdirectory with specifications above.
spatPlot(my_gobject, 
         cell_color = 'ClusterName', 
         save_plot = TRUE,
         return_plot = FALSE,
         save_param = list(save_folder = results_directory, # Create subdirectory
                           save_name = 'my_name', 
                           save_format = 'png', 
                           units = 'in',
                           base_height = 9,
                           base_width = 9))
```


## Just View the Plot
See [Giotto Object](./getting_started_gobject.html) for more details.  
Set both *save_plot* and *return_plot* to FALSE.
```{r eval=TRUE}
# Plot without saving
spatPlot(my_gobject, 
         cell_color = 'ClusterName', 
         save_plot = FALSE, return_plot = FALSE, show_plot = T)
```


## Just save the plot (FASTEST for large datasets!)
See [Giotto Object](./getting_started_gobject.html) for more details.  
Set show_plot and return_plot to FALSE, set save_plot to TRUE.
```{r eval=TRUE}
spatPlot(my_gobject, 
         cell_color = 'ClusterName', 
         save_plot = TRUE, return_plot = FALSE, show_plot = FALSE,
         save_param = list(save_name = 'only_save'))
```

## Session Info
```{r}
sessionInfo()
```
